---
title: "mfe: Meta-Feature Extractor"
author: "Adriano Rivolli, Luis P. F. Garcia and Andre C. P. L. F. de Carvalho"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{mfe: Meta-Feature Extractor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Extracts meta-features from datasets to support the design of recommendation 
systems based on Meta-Learning (MtL). The meta-features, also called 
characterization measures, are able to characterize the complexity of datasets 
and to provide estimates of algorithm performance. The package contains not 
only the standard, but also more recent characterization measures. By making 
available a large set of meta-feature extraction functions, this package allows 
a comprehensive data characterization, a deep data exploration and a large 
number of MtL-based data analysis.

In MtL, meta-features are designed to extract general properties able to 
characterize datasets. The meta-feature values should provide relevant 
evidences about the performance of algorithms, allowing the design of MtL-based 
recommendation systems. Thus, these measures must be able to predict, with a 
low computational cost, the performance of the algorithms under evaluation. In 
this package, the meta-feature measures are divided into five groups:

* **General**: General information related to the dataset, also known as simple 
measures, such as number of instances, attributes and classes.
* **Statistical**: Standard statistical measures to describe the numerical 
properties of a distribution of data and also the discriminant measures.
* **Information-theoretic**: Particularly appropriate to describe discrete 
(categorical) attributes and their relationship with the classes.
* **Decision Tree Model-based**: Measures designed to extract characteristics 
like the depth, the shape and size of a Decision Tree (DT) model induced from a 
dataset.
* **Landmarking**: Represents the performance of simple and efficient learning 
algorithms.

In the following sections we will briefly introduce how to use the `mfe` 
package to extract all the measures using standard methods as well as to extract 
specific measures using methods for each group. Once the package is loaded, the 
vignette is also available inside R with the command `browseVignettes`.

# Extracting meta-features

The standard way to extract meta-features is using the `metafeatures` methods. 
The method can be used by a symbolic description of the model (formula) or by a 
data frame. The parameters are the dataset and the group of measures to be 
extracted. By default, the method extract all the measures. For instance:

```{r}
library(mfe)

## Extract all measures using formula
iris.info <- metafeatures(Species ~ ., iris)

## Extract all measures using data frame
iris.info <- metafeatures(iris[,1:4], iris[,5])

## Extract general, statistical and information-theoretic measures
iris.info <- metafeatures(Species ~ ., iris, 
                          groups=c("general", "statistical", "infotheo"))
```

Several measures return more than one value. To aggregate them, post processing 
methods can be used. It is possible to compute min, max, mean, median, kurtosis,
 standard deviation, among others. The default methods are the `mean` and the 
 `sd`. For instance:

```{r}
## Compute all measures using min, median and max 
iris.info <- metafeatures(Species ~ ., iris, summary=c("min", "median", "max"))
                          
## Compute all measures using quantile
iris.info <- metafeatures(Species ~ ., iris, summary="quantile")
```

# Extracting meta-features by group

To customize the measure extraction, is necessary to use specific methods for 
each group of measures. For instance, `infotheo` and `statistical` compute 
the information theoretical and the statistical measures, respectively. The 
following examples illustrate these cases:

```{r}
## Extract two information theoretical measures
stat.iris <- infotheo(Species ~ ., iris, 
                         features=c("attrEnt", "jointEnt"))

## Extract three statistical measures
disc.iris <- statistical(Species ~ ., iris, 
                            features=c("cancor", "cor", "iqr"))

## Extract the histogram for the correlation measure
hist.iris <- statistical(Species ~ ., iris, 
                            features="cor", summary="hist")
```

Different from the `metafeatures` method, these methods receive a parameter 
called `features`, to define which features are required, and return a list 
instead of a numeric vector. In additional, some groups can be customized using
additional arguments.

There are five measure groups which can be either general information about the 
dataset, statistical information, descriptors about information theoretical, 
measures designed to extract characteristics about the DT model or landmarks 
which represent the performance of simple algorithms applied to the dataset. 
The following example show the available groups:

```{r}
## Show the the available groups
ls.metafeatures()
```

## General

These are the most simple measures for extracting general properties of the 
datasets. For instance, `nrAttr` and `nrClass` are the total number of 
attributes in the dataset and the number of output values (classes) in the 
dataset, respectively. To list the measures of this group use `ls.general()`. 
The following examples illustrate these measures:

```{r}
## Show the the available general measures
ls.general()

## Extract all general measures
general.iris <- general(Species ~ ., iris)

## Extract two general measures
general(Species ~ ., iris, features=c("nrAttr", "nrClass"))
```

The general measures return a list named by the requested measures. The 
`post.processing` methods are applied only for the `freqClass` meta-feature. For
instance, to extract the minimum, maximum and the standard deviation of the
classes proportion use:

```{r}
## Extract two general measures
general(Species ~ ., iris, features="freqClass", summary=c("min", "max", "sd"))
```

## Statistical

Statistical meta-features are the standard statistical measures to describe the 
numerical properties of a distribution of data. As it requires only numerical 
attributes, the categorical data are transformed to numerical. For instance, 
`cor` and `skewness` are the absolute correlation between of each pair of 
attributes and the skewness of the numeric attributes in the dataset, 
respectively. To list the measures of this group use `ls.statistical()`. The 
following examples illustrate these measures:

```{r}
## Show the the available statistical measures
ls.statistical()

## Extract all statistical measures
stat.iris <- statistical(Species ~ ., iris)

## Extract two statistical measures
statistical(Species ~ ., iris, features=c("cor", "skewness"))
```

The statistical group can use two additional parameter called `by.class` and 
`transform`. To the former the default is `by.class=FALSE` which means that the 
meta-features are computed without consider the classes values. Otherwise, 
the measure is extracted using the instances separated by class. In the latter, 
the default value is `transform=TRUE` which means that categorical attributes 
will be transformed to numeric. The following example shows the use of these 
two definitions:

```{r}
## Extract correlation using instances by classes
statistical(Species ~ ., iris, features="cor", by.class=TRUE)

## Ignore the class attributes
aux <- cbind(class=iris$Species, iris)
statistical(Species ~ ., aux, transform=FALSE)
```

Note that, in the first example the values and the cardinality of the measure 
are different since the correlation between the attributes were computed using 
the instances for each class separately. The `post.processing` methods are 
applied in these measures since they return multiple values. To define which 
them should be applied use the `summary` parameter, as detailed in the 
`post.processing` method.

## Information Theoretical

Information theoretical meta-features are particularly appropriate to describe
discrete (categorical) attributes, but they also fit continuous ones using a 
discretization process. These measures are based on information theory. For 
instance, `normClassEnt` and `mutInf` are the normalized entropy of the class 
and the common information shared between each attribute and the class in the 
dataset, respectively. To list the measures of this group use `ls.infotheo()`. 
The following examples illustrate these measures:

```{r}
## Show the the available information theoretical measures
ls.infotheo()

## Extract all information theoretical measures
inf.iris <- infotheo(Species ~ ., iris)

## Extract two information theoretical measures
infotheo(Species ~ ., iris, features=c("normClassEnt", "mutInf"))
```

The Information theoretical group can use one additional parameter called 
`transform`. Using the default value `transform=TRUE` the continuous attributes 
will be discretized. The following example shows the use of this definition:

```{r}
## Ignore the discretization process
aux <- cbind(class=iris$Species, iris)
infotheo(Species ~ ., aux, transform=FALSE)
```

The information theoretical measures return a list named by the requested 
measures. The `post.processing` methods are applied in some measures since they 
return multiple values. To define which them should be applied use the 
`summary` parameter, as detailed in the section **Post Processing Methods**.

## Model based

These measures describe characteristics of the investigated models. These 
meta-features can include, for example, the description of the DT induced for a 
dataset, like its number of leaves (`leaves`) and the number of nodes (`nodes`) 
of the tree. The following examples illustrate these measures:

```{r}
## Show the the available model.based measures
ls.model.based()

## Extract all model.based measures
land.iris <- model.based(Species ~ ., iris)

## Extract three model.based measures
model.based(Species ~ ., iris, features=c("leaves", "nodes"))
```

The DT model based measures return a list named by the requested measures. The 
`post.processing` methods are applied in these measures since they return 
multiple values. To define which them should be applied use the `summary` 
parameter, as detailed in the `post.processing` method.

## Landmarking

Landmarking measures are simple and fast algorithms, from which performance 
characteristics can be extracted. These measures include the performance of 
simple and efficient learning algorithms like Naive Bayes (`naiveBayes`) and 
1-Nearest Neighbor (`oneNN`). The following examples illustrate these measures:

```{r}
## Show the the available landmarking measures
ls.landmarking()

## Extract all landmarking measures
land.iris <- landmarking(Species ~ ., iris)

## Extract two landmarking measures
landmarking(Species ~ ., iris, features=c("naiveBayes", "oneNN"))
```

The performance extraction of these measures without a cross validation step can
 cause model overfitting in the data. Therefore the `landmarking` function 
has the parameter `folds` to define the number of `k`-fold cross-validation and
the parameter `score` to select the performance measure. The following example 
show how to set this value:

```{r}
## Extract one landmarking measures with folds=2
landmarking(Species ~ ., iris, features="naiveBayes", folds=2)

## Extract one landmarking measures with folds=2
landmarking(Species ~ ., iris, features="naiveBayes", score="kappa")
```

The landmarking measures return a list named by the requested measures. The 
`post.processing` methods are applied in these measures since they return 
multiple values. To define which them should be applied use the `summary` 
parameter, as detailed in the `post.processing` method.

## Relative

The relative group is the landmarking with sampling and ranking strategies. The 
sampling strategy decreases the computational cost of the landmarking by 
selecting a subsample of the original examples. The ranking strategy capture 
relative information between the performance of the algorithms. The following 
examples illustrate these measures:

```{r}
## Show the the available relative measures
ls.relative()

## Extract all relative measures
real.iris <- relative(Species ~ ., iris)

## Extract all relative measures with half of the samples
relative(Species ~ ., iris, size=0.5)

## Extract two relative measures
relative(Species ~ ., iris, features=c("naiveBayes", "oneNN"))
```

## Clustering

Clustering measures extract information about dataset based on external 
validation indexes. The main ideia is measure the complexity of the dataset 
using indexes able to check information about the predictive attributes and the 
label. The following examples illustrate these measures:

```{r}
## Show the the available clustering measures
ls.clustering()

## Extract all clustering measures
clus.iris <- clustering(Species ~ ., iris)

## Extract two clustering measures
clustering(Species ~ ., iris, features=c("vdu", "vdb"))
```

# Post Processing Methods

Several meta-features generate multiple values and `mean` and `sd` are the 
standard method to summary these values. In order to increase the flexibility, 
the `mfe` package implemented the post processing methods to deal with multiple 
measures values. This method is able to deal with descriptive statistic 
(resulting in a single value) or a distribution (resulting in multiple values).

The post processing methods are setted using the parameter `summary`. It is 
possible to compute min, max, mean, median, kurtosis, standard deviation, among 
others. Any R method, can be used, as illustrated in the following examples:

```{r}
## Apply several statistical measures as post processing
statistical(Species ~ ., iris, "cor", 
               summary=c("kurtosis", "max", "mean", "median", "min", "sd", 
                         "skewness", "var"))

## Apply quantile as post processing method
statistical(Species ~ ., iris, "cor", summary="quantile")

## Get the default values without summarize them
statistical(Species ~ ., iris, "cor", summary=c())
```

Beyond these R default methods, two additional post processing methods are 
available in the `mfe` package: `hist` and `non.aggregated`. The first computes 
a histogram of the values and returns the frequencies of in each bins. The 
extra parameters `bins` can be used to define the number of values to be 
returned. The parameters `min` and `max` are used to define the range of the 
data. The second is a way to obtain all values from the measure and has the 
same effect of the use of an empty list. The following code illustrate examples 
of the use of these post processing methods:

```{r}
## Apply histogram as post processing method
statistical(Species ~ ., iris, "cor", summary="hist")

## Apply histogram as post processing method and customize it
statistical(Species ~ ., iris, "cor", summary="hist", bins=5, min=0, max=1)

## Extract all correlation values
statistical(Species ~ ., iris, "cor", summary="non.aggregated")
```

It is also possible define an user's post processing method, like this:

```{r}
## Compute the absolute difference between the mean and the median 
my.method <- function(x, ...) abs(mean(x) - median(x))

## Using the user defined post processing method
statistical(Species ~ ., iris, "cor", summary="my.method")
```

# Summary

In this paper the `mfe` package, aimed to extract meta-features from dataset, 
has been introduced. The functions supplied by this package allow both their 
use in MtL experiments as well as perform data analysis using characterization 
measures able to describe datasets. Currently, six groups of meta-features can 
be extracted for any classification dataset. These groups and features represent
the standard and the state of the art characterization measures. 

The `mfe` package was designed to be easily customized and extensible. The 
development of the `mfe` package will continue in the near future by including 
new meta-features, group of measures supporting regression problems and MtL 
evaluation measures.
