---
title: "mfe: Meta-Features Extractor for Meta-Learning"
author: "Adriano Rivolli and Luis Paulo F. Garcia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The mfe package is designed to extract meta-features from datasets. The 
meta-features can be understood as characterization measures able to describe 
datasets to support recommendation systems based on Meta-learning (MtL). The 
package contains the standard and the state of the art characterization measures
 with the goal to improve the MtL experiments and also guide the complexity 
dataset understanding. In the current version, only classification datasets are 
supported, however, we plan support regression datasets in the future versions. 

The meta-features are designed to extract general properties of datasets and 
provide evidences about the performance of algorithms in MtL recomendation 
systems. These measures must be able to predict, with a low computational cost, 
the performance of these algorithms. The measures used in MtL can be divided 
into six groups:

* **General**: General information related to the dataset, also known as simple 
measures, such as number of instances, attributes and classes.
* **Statistical**: Standard statistical measures to describe the numerical 
properties of a distribution of data.
* **Discriminant**: Measures computed using the discriminant analysis.
* **Information-theoretic**: Particularly appropriate to describe discrete 
(categorical) attributes and thei relationship with the classes.
* **Decision Tree Model-based**: Measures desined to extract characteristics 
like the depth, the shape and size of a Decision Tree model induced from a 
dataset.
* **Landmarking**: Represents the performance of some simple and efficient 
learning algorithms.

# Extracting meta-features

The simplest way to extract meta-features is using the `metafeatures` method. 
The method can be usage by a symbolic description of the model or by a data 
frame. The parameters are the dataset and the group of measures to be extracted.
 To extract all the measures, the parameter "group" needs to be set as "all". 
 For instance:

```{r}
library(mfe)
data("iris")

## Extract all measures using formula
iris.info <- metafeatures(Species ~ ., iris, groups="all")

## Extract all measures using data frame
iris.info <- metafeatures(iris[1:4], iris[,5], groups="all")

## Extract general, statistical and information-theoretic measures
iris.info <- metafeatures(Species ~ ., iris, 
                          groups=c("general", "statistical", "infotheo"))

## Show the the available groups
ls.metafeatures()
```

Several measures return more than one value. To agregate them, post processed 
methods can be used. It is possible to compute min, max, mean, median, kurtosis,
 standard deviation, among others (see the `post.processing` documentation for 
 more details). The default methods are the `mean` and the `sd`. For instance:


```{r}
## Compute all measures using min, median and max 
iris.info <- metafeatures(Species ~ ., iris, summary=c("min", "median", "max"))
                          
## Compute all measures using quantile
iris.info <- metafeatures(Species ~ ., iris, summary="quantile")
```

## Extracting meta-features by group

To customize the measure extraction, is necessary to use specific methods for 
each group of measures. For instance, `mf.general` and `mf.statistical` compute 
the general and the statistical measures, respectively. To list the measures of 
these groups use `ls.general()` and `ls.statistical()`. The folloing examples 
illustrate this cases:

```{r}
## Extract two statistical measures
stat.iris <- mf.statistical(Species ~ ., iris, 
                            features=c("correlation", "variance"))

## Extract two discriminant measures
disc.iris <- mf.discriminant(Species ~ ., iris, 
                             features=c("cancor", "cancor.fract"))

## Extract the histogram for the correlation measure
hist.iris <- mf.statistical(Species ~ ., iris, 
                            features="correlation", summary="hist")
```

Different from the `metafeatures` method, these methods return a list instead of
 a numeric vector. To get all measures without post processing, use 
 `summary=non.aggregated` like this:

```{r}
## Extract all correlation values
cor.iris <- mf.statistical(Species ~ ., iris, 
                           features="correlation", summary="non.aggregated", 
                           by.class=FALSE)
```

In the next sections, each group is detailed and the post processing methods are
 described.

# Meta-features Groups

## General

## Statistical

## Discriminant

## Information Theoretical

## Model based

## Landmarking

# Post Processing Methods


# Summary
