---
title: "mfe: Meta-Features Extraction for Meta-Learning"
author: "Adriano Rivolli and Luis Paulo F. Garcia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Overview

The simplest way to extract meta-features is using the `metafeatures` method.
You just need to inform which group of measures do you need or use "all" for 
all measures. For instance:
```{r}
library(mfe)
data("iris")

## Extract general, statistical and information-theoretic measures
iris.info <- metafeatures(Species ~ ., iris, c("general", "statistical", "infotheo"))

## Extract all meta-features
iris.info <- metafeatures(iris[1:4], iris$Species)

## List the available groups
ls.metafeatures()
```

Several measures generates more than one values and them can be post processing
using distinct alternatives. It is possible compute the min, max, mean, median,
kurtosis, standard deviation, among others (See the `post.processing` 
documentation for more details). The default is the `mean` and the `sd` however
you can use anothers. For instance:
```{r}
## Summarize multiples meta-features values using min, median and max 
iris.info <- metafeatures(Species ~ ., iris, "statistical", 
                          summary=c("min", "median", "max"))
                          
## Summarize multiples meta-features values using quantile
iris.info <- metafeatures(Species ~ ., iris, "statistical", summary="quantile")
```

Despite the simplicity it is not possible customize the meta-features, for this 
is necessary to use the specific methods. There is a method for each group of 
measure, for instance `mf.general` and `mf.statistical` compute the general and
the statistical measures respectively. To list the measures of these groups use 
`ls.general()` and `ls.statistical()`. The folloing examples illustrate this 
cases:
```{r}
## Extract some statistical meta-features
stat.iris <- mf.statistical(iris[1:4], iris[5], c("correlation", "variance"))

## Extract the histogram of correlation
hist.iris <- mf.statistical(iris[1:4], iris[5], "correlation", "hist")

## Extract some discriminant meta-features
disc.iris <- mf.discriminant(iris[1:4], iris[5], c("cancor", "cancor.fract"))
```

Different from the `metafeatures` method, these methods return a list instead of
a numeric vector. To get all meta-feature values without post processing use `summary=non.aggregated` like this:
```{r}
## Extract all correlation values
cor.iris <- mf.statistical(iris[1:4], iris[5], "correlation", 
                           summary="non.aggregated", by.class=FALSE)
```
