---
title: "mfe: Meta-Features Extraction for Meta-Learning"
author: "Adriano Rivolli and Luis Paulo F. Garcia"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

The mfe package is designed to extract meta-features from datasets. 
The meta-features can be understood as characterization measures able to 
describe datasets to support recommendation systems based on meta-learning.
The package contains traditional and new characterization measures with the 
goal to improve the meta-learning experiments and also guide the complexity 
dataset understanding.

The following groups of measures are available:

* **General** - General information related to the dataset, also known as simple 
measures, such as number of instances, attributes and classes.
* **Statistical** - Standard statistical measures to describe the numerical 
properties of a distribution of data.
* **Discriminant** - Measures computed using the discriminant analysis.
* **Information-theoretic** - Particularly appropriate to describe discrete 
(categorical) attributes and thei relationship with the classes.
* **Model-based** - Measures desined to extract characteristics like the depth, 
the shape and size of a Decision Tree model induced from a dataset.
* **Landmarking** - Represents the performance of some simple and efficient 
learning algorithms.

# Getting Started

## Installation
The installation process is similar to other packages available on CRAN:
```r
install.packages("mfe")
```

It is possible to install the development version using
```r
devtools::install_github("rivolli/mfe")
```

## Extracting meta-features
The simplest way to extract meta-features is using the `metafeatures` method.
It receives as parameter which groups of measures will be needed. To get all 
groups, use the value "all". 
For instance:
```{r}
library(mfe)
data("iris")

## Extract general, statistical and information-theoretic measures
iris.info <- metafeatures(Species ~ ., iris, c("general", "statistical", "infotheo"))

## Extract all meta-features
iris.info <- metafeatures(iris[1:4], iris$Species, "all")

## List the available groups
ls.metafeatures()
```

Several measures generates more than one values and them can be post processing
using distinct alternatives. It is possible compute the min, max, mean, median,
kurtosis, standard deviation, among others (Described in the section Post Processing). 
The default is the `mean` and the `sd` however any other method can be used. 
For instance:
```{r}
## Summarize multiples meta-features values using min, median and max 
iris.info <- metafeatures(Species ~ ., iris, "statistical", 
                          summary=c("min", "median", "max"))
                          
## Summarize multiples meta-features values using quantile
iris.info <- metafeatures(Species ~ ., iris, "statistical", summary="quantile")
```

## Meta-features by group
Despite the simplicity of the method `metafeatures`, it is not possible choose 
or customize the meta-features. For this is necessary to use the group methods, 
they begin with `mf.` prefix, for instance `mf.general` and `mf.statistical` 
compute the general and the statistical measures respectively. 
To list the measures of these groups use `ls.general()` and `ls.statistical()`. 
The folloing examples illustrate this cases:
```{r}
## Extract some statistical meta-features
stat.iris <- mf.statistical(iris[1:4], iris[5], c("correlation", "variance"))

## Extract the histogram of correlation
hist.iris <- mf.statistical(iris[1:4], iris[5], "correlation", "hist")

## Extract some discriminant meta-features
disc.iris <- mf.discriminant(iris[1:4], iris[5], c("cancor", "cancor.fract"))
```

Different from the `metafeatures` method, these methods return a list instead of
a numeric vector. To get all meta-feature values without post processing use `summary=non.aggregated` like this:
```{r}
## Extract all correlation values
cor.iris <- mf.statistical(iris[1:4], iris[5], "correlation", 
                           summary="non.aggregated", by.class=FALSE)
```

In the next sections, each group is detailed and the post processing methods are described.

# Meta-features Groups

## General

## Statistical

## Discriminant

## Information Theoretical

## Model based

## Landmarking

# Post Processing Methods


# Summary
